Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{ I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback. }

### 2023-06-27 paper/paper.tex
#32 remove blanks (compiles in 7607)

A key economic assumption is needed, even before the econometric ones, for these interaction terms to tell us anything about state power: the power of the state owning city $i$ is correlated with the manner in which the state acquired city $i$. In particular, I assume that conquests are more strongly positively correlated with state power than successions. Support for this assumption comes from \cite{levine2021}'s model of conflict-driven evolution. BLANK. I will discuss its validity at length later on.

If this assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of cities acquired by succession. Note that successions are the omitted category --- there is no interaction term between the post-treatment indicator and $\mathbf{1}(Succession_{it})$ in Equation \eqref{eq:baseline} --- so that the coefficient $\theta_{conquest}$ captures exactly what I want to estimate: the difference in trends after conquests vis-a-vis successions.

The econometric assumption that must hold true for $\theta_{conquest}$ to be a causal effect is a different version of the parallel trends assumption. In particular, it must hold that, had a conquered city $i$ counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession. BLANK.

@@

The only subset of construction for which the difference in trends is indeed significant is state construction --- palaces, castles, military and administrative buildings. We can see in Column (2) that these are 5 percentage points \textit{less} likely to be built in a city after it was conquered; the effect is significant at the 5\% level. BLANK.

@@

Together, Table \ref{tab:controls_50y} and Figure \ref{fig:conflict_map} have the interesting implication of spatially heterogeneous differences in treatment effects. The null difference in Column (1) could be an average of the negative difference in the southwest and a hypothetical positive difference in the northeast. Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory: the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
{
Of course, England and Sweden would attack via the North and Baltic seas, but it seems a reasonable assumption that competition for territory would be more intense between states that share a land border.
} 
BLANK \cite{cervellati2022}. I leave this for future research to explore.

>>>>

A key economic assumption is needed, even before the econometric ones, for these interaction terms to tell us anything about state power: the power of the state owning city $i$ is correlated with the manner in which the state acquired city $i$. In particular, I assume that conquests are more strongly positively correlated with state power than successions. The idea is that conquering a city means winning it in a conflict --- which is easier if you are a powerful state --- while inheriting it only requires that you, or even your distant ancestors, entered a fortunate marriage. I will discuss the validity of this assumption at length later on.

If the assumption holds true, the effect of state power should manifest itself in the difference between the trend of cities acquired by conquest and the trend of cities acquired by succession. Note that successions are the omitted category --- there is no interaction term between the post-treatment indicator and $\mathbf{1}(Succession_{it})$ in Equation \eqref{eq:baseline} --- so that the coefficient $\theta_{conquest}$ captures exactly what I want to estimate: the difference in trends after conquests vis-a-vis successions.

The econometric assumption that must hold true for $\theta_{conquest}$ to be a causal effect is a different version of the parallel trends assumption. In particular, it must hold that, had a conquered city $i$ counterfactually been acquired by succession rather than conquest, it would have trended the same as the other cities acquired by succession.

@@

The only subset of construction for which the difference in treatment effects is indeed significant is state construction --- palaces, castles, military and administrative buildings. We can see in Column (2) that these are 5 percentage points \textit{less} likely to be built in a city after it was conquered; the effect is significant at the 5\% level. It is hard to explain this fact with state power; however, keeping in mind that this is the effect of conquests relative to successions, it could be that rulers who inherit a city feel a bigger desire to impress and make their presence known to the townspeople than someone who just marched an army into it.

@@

Together, Table \ref{tab:controls_50y} and Figure \ref{fig:conflict_map} have the interesting implication of spatially heterogeneous differences in treatment effects. The null difference in Column (1) could be an average of the negative difference in the southwest and a hypothetical positive difference in the northeast. Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory: the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
{
Of course, England and Sweden would attack via the North and Baltic seas, but it seems a reasonable assumption that competition for territory would be more intense between states that share a land border.
} 
However, I leave this topic for future research to explore.
